# Supplementary exercises {#sec-ex}

```{r visibility, echo = FALSE}
vis <- FALSE
```

## Raw data and identification results


- Download the 3 first mzML and mzID files from the
  [PXD022816](https://www.ebi.ac.uk/pride/archive/projects/PXD022816)
  project [@Morgenstern:2020].


```{r get_PXD022816, echo = vis, include = vis}
## Getting data from PX/PRIDE
library(rpx)

## https://www.ebi.ac.uk/pride/archive/projects/PXD022816
## RawBeans: A Simple, Vendor-Independent, Raw-Data Quality-Control
## Tool ()

PXD022816 <- PXDataset("PXD022816")
PXD022816

pxfiles(PXD022816)

(mzids <- pxget(PXD022816, grep("mzID", pxfiles(PXD022816))[1:3]))
(mzmls <- pxget(PXD022816, grep("mzML", pxfiles(PXD022816))[1:3]))
```


- Generate a `Spectra` object and a table of filtered PSMs. Visualise
  the total ion chromatograms and check the quality of the
  identification data by comparing the density of the decoy and target
  PSMs id scores for each file.

```{r spectra_PXD022816, echo = vis, include = vis}
## Loading raw data
library("Spectra")
sp <- Spectra(mzmls)
sp

## number of spectra per file
table(basename(sp$dataOrigin))
## all levels are centroided
table(sp$centroided, sp$msLevel)
```

```{r mslevel_PXD022816, message = FALSE, fig.width = 12, echo = vis, include = vis}
library("ggplot2")
library("tidyr")
library("magrittr")

## Chromatograms
filterMsLevel(sp, 1) |>
    spectraData() |>
    as_tibble() |>
    ggplot(aes(x = rtime,
               y = totIonCurrent,
               colour = basename(dataOrigin))) +
    geom_line()
```

```{r is_PXD022816, message = FALSE, fig.width = 12, echo = vis, include = vis}
## Identification data
library("PSMatch")
id <- PSM(mzids)

## Number of PSMs per acquisition
table(id$idFile)

tidyr::as_tibble(id) |>
    ggplot(aes(x = MetaMorpheus.score,
               colour = isDecoy)) +
    geom_density() +
    facet_wrap(~ spectrumFile)
```

```{r idfilter_PXD022816, echo = vis, include = vis}
id_filtered <- filterPSMs(id)
max(id_filtered$PSM.level.q.value)
```



- Join the raw and identification data. Beware though that the joining
  must now be performed by spectrum ids and by files.

```{r joinsp_PXD022816, echo = vis, include = vis}
## primary key for spectra
sp$pkey <-
    paste0(sub("^.+_QEP", "QEP", basename(dataOrigin(sp))),
           gsub("^.+=", "::", sp$spectrumId))
head(sp$pkey)

## primary key for PSMs
id_filtered$pkey <-
    paste0(gsub("^.+\\QEP", "QEP", id_filtered$spectrumFile),
           sub("^.+=", "::", id_filtered$spectrumID))
head(id_filtered$pkey)

## For simplicity, let's keep single hits per spectrum id.
## Alternatively, explore duplicates and use QFeatures::reduceDataFrame
id_filtered <- id_filtered[!duplicated(id_filtered$pkey), ]

head(id_filtered$pkey)

sp <- joinSpectraData(sp, id_filtered, by.x = "pkey")

## Number of MS2 scans with a PSM
table(!is.na(filterMsLevel(sp, 2)$sequence))
```

- Extract the PSMs that have been matched to peptides from protein
  `O43175` and compare and cluster the scans. Hint: once you have
  created the smaller `Spectra` object with the scans of interest,
  switch to an in-memory backend to seed up the calculations.


```{r compare_PXD022816, echo = vis, include = vis}
sp_O43175 <- sp[which(sp$DatabaseAccess == "O43175")]
sp_O43175 <- setBackend(sp_O43175, MsBackendDataFrame())
sp_O43175

cmat <- compareSpectra(sp_O43175)
rownames(cmat) <-
    colnames(cmat) <- strtrim(sp_O43175$sequence, 3)

pheatmap::pheatmap(cmat)

(i <- which(rownames(cmat) == "DLP"))

plotSpectra(sp_O43175[i], labels = labelFragments,
            labelPos = 3, labelCol = "steelblue",
            main = sp_O43175$sequence[i])

spectraData(sp_O43175[i])$precursorCharge
spectraData(sp_O43175[i])$precursorMz
spectraData(sp_O43175[i])$modName
```

```{r plotmirror_PXD022816, echo = vis, include = vis}
## Directly compare spectra with/without Carbamyl
plotSpectraMirror(sp_O43175[4], sp_O43175[9])
plotSpectraMirror(sp_O43175[2], sp_O43175[10])
```


- Generate total ion chromatograms for each acquisition and annotate
  the MS1 scans with the number of PSMs using the
  `countIdentifications()` function, as shown above. The function will
  automatically perform the counts in parallel for each acquisition.

```{r countIds, echo = vis, include = vis}
sp <- countIdentifications(sp)
table(msLevel(sp), sp$countIdentifications)
```

```{r tic_PXD022816, echo = vis, include = vis}
sp |>
 filterMsLevel(1) |>
 spectraData() |>
 as_tibble() |>
 ggplot(aes(x = rtime,
            y = totIonCurrent)) +
     geom_line(alpha = 0.25) +
     geom_point(aes(colour = ifelse(countIdentifications == 0,
                                    NA, countIdentifications)),
                size = 0.75,
                alpha = 0.5) +
     scale_colour_gradient(low = "orange", high = "red") +
     facet_grid(sub("^.+_", "", basename(dataOrigin)) ~ .) +
     labs(colour = "Number of ids")
```

## Search engine

## Quantitative data processing

Following up from the quantitative data analysis seen on chapter
\@ref(sec-quant), the following file includes a third condition C and
a two additional lab, tallying now 27 samples.


```{r cptac_abc, message = FALSE}
f <- MsDataHub::cptac_a_b_c_peptides.txt()
```

```{r, echo = FALSE, message = FALSE}
library(QFeatures)
i <- grep("Intensity\\.", names(read.delim(f)))
cptac2 <- readSummarizedExperiment(f, quantCols = i,
                                   sep = "\t",
                                   fnames = "Sequence")
colnames(cptac2) <- sub("I.+\\.", "", colnames(cptac2))
cptac2$id <- sub("6.+_", "", colnames(cptac2))
cptac2$condition <- sub("_[1-9]", "", colnames(cptac2))
cptac2$lab <- rep(rep(c("LTQ-Orbitrap_86",
                       "LTQ-OrbitrapO_65",
                       "LTQ-OrbitrapW_56"), each = 3),
                  3)
knitr::kable(table(colData(cptac2)[, -1]))
```

The full design is shown below.

```{r, echo=FALSE}
cd <- data.frame(colData(cptac2))
cd$previous <- "new"
cd[colnames(cptac_se), "previous"] <- ""
knitr::kable(cd, rownames = TRUE)
```

- Repeat the analysis described in chapter \@ref(sec-quant) using the
  extended dataset, trying to optimise true positive results and
  avoiding false positive. Think about the best experimental design
  approach, how to best process the data, visualising important steps
  along the way, to conclude with a volcano plot and a table tallying
  the number of true/false positive/negative results.