# Annex {#sec-anx}


## Raw MS data under the hood: the `mzR` package {#sec-raw2}

The `mzR` package in a direct interface to the
[proteowizard](http://proteowizard.sourceforge.net/) code base. It
includes a substantial proportion of *pwiz*'s C/C++ code for fast and
efficient parsing of these large raw data files.

Let's start by using some raw data files from the `msdata`
package. After loading it, we use the `proteomics()` function to
return the full file names for two raw data files. We will start by
focusing on the second one.

```{r rwfiles}
f <- msdata::proteomics(full.names = TRUE)
f
(f2 <- grep("20141210", f, value = TRUE))
```

The three main functions of `mzR` are

* `openMSfile` to create a file handle to a raw data file
* `header` to extract metadata about the spectra contained in the file
* `peaks` to extract one or multiple spectra of interest.

Other functions such as `instrumentInfo`, or `runInfo` can be used to
gather general information about a run.


```{r rawms}
library("mzR")
ms <- openMSfile(f2)
ms
```

```{r hd}
hd <- header(ms)
dim(hd)
names(hd)
```

```{r peaks}
head(peaks(ms, 117))
str(peaks(ms, 1:5))
```

`r msmbstyle::question_begin()`

Let's extract the index of the MS2 spectrum with the highest base peak
intensity and plot its spectrum. Is the data centroided or in profile
mode?
`r msmbstyle::question_end()`


`r msmbstyle::solution_begin()`

```{r ex_raw, echo=TRUE, eval=TRUE, fig.align='center'}
hd2 <- hd[hd$msLevel == 2, ]
i <- which.max(hd2$basePeakIntensity)
hd2[i, ]
pi <- peaks(ms, hd2[i, 1])
plot(pi, type = "h")
mz <- hd2[i, "basePeakMZ"]
plot(pi, type = "h", xlim = c(mz - 0.5, mz + 0.5))
```

`r msmbstyle::solution_end()`


`r msmbstyle::question_begin()`

Pick an MS1 spectrum and visually check whether it is centroided or in
profile mode.

`r msmbstyle::question_end()`


`r msmbstyle::solution_begin()`

```{r ex_raw2}
## Zooming into spectrum 300 (an MS1 spectrum).
j <- 300
pj <- peaks(ms, j)
plot(pj, type = "l")
mz <- hd[j, "basePeakMZ"]
plot(pj, type = "l", xlim = c(mz - 0.5, mz + 0.5))
```

`r msmbstyle::solution_end()`


## PSM data under the hood {#sec-id2}

There are two packages that can be used to parse `mzIdentML` files,
namely `mzR` (that we have already used for raw data) and `mzID`. The
major difference is that the former leverages C++ code from
`proteowizard` and is hence faster than the latter (which uses the
`XML` R package). They both work in similar ways.

```{r idpkgs, echo = FALSE, eval = FALSE}
x <- RforProteomics:::msDataTab()
sel <- x[, 1] %in% c("Identification")
knitr::kable(x[sel, ])
```

```
|Data type      |File format |Data structure |Package |
|:--------------|:-----------|:--------------|:-------|
|Identification |mzIdentML   |mzRident       |mzR     |
|Identification |mzIdentML   |mzID           |mzID    |
```

Which of these packages is used by `PSM()` can be defined by the
`parser` argument, as documented in `?PSM`.

### `mzID` {-}

The main functions are `mzID` to read the data into a dedicated data
class and `flatten` to transform it into a `data.frame`.

```{r mzid1, warning = FALSE}
idf
library("mzID")
id <- mzID(idf)
id
```

Various data can be extracted from the `mzID` object, using one the
accessor functions such as `database`, `software`, `scans`, `peptides`,
... The object can also be converted into a `data.frame` using the
`flatten` function.

```{r flatmzid}
head(flatten(id))
```

### `mzR` {-}

The `mzR` interface provides a similar interface. It is however much
faster as it does not read all the data into memory and only extracts
relevant data on demand. It has also accessor functions such as
`softwareInfo`, `mzidInfo`, ... (use `showMethods(classes = "mzRident", where = "package:mzR")`)
to see all available methods.

```{r idmzr}
library("mzR")
id2 <- openIDfile(idf)
id2
softwareInfo(id2)
```

The identification data can be accessed as a `data.frame` with the
`psms` accessor.

```{r psmsmzr}
head(psms(id2))
```
