# References and session information {#sec-si}

## Additional materials

- The [Single-cell proteomics data analysis using `QFeatures` and
  `scp`](https://lgatto.github.io/QFeaturesScpWorkshop2021/) workshop
  is provided as two vignettes. The first one provides a general
  introduction to the `QFeatures` class in the general context of mass
  spectrometry-based proteomics data manipulation. The second vignette
  focuses on single-cell application and introduces the `scp` package
  [@Vanderaa:2021] as an extension of `QFeatures`. This second
  vignette also provides exercises that give the attendee the
  opportunity to apply the learned concepts to reproduce a published
  analysis on a subset of a real data set.

- The [Seamless Integration of Mass Spectrometry Data from Different
  Sources](https://jorainer.github.io/SpectraTutorials/) workshop
  explains the `Spectra` package and shows how this new infrastructure
  can be used to represent and analyze Mass Spectrometry (MS)
  data. The use of different data type and format specific backends
  guarantees an easy expandability of the infrastructure. This is
  illustrated in this tutorial on a simple use case, in which MS/MS
  (a.k.a. MS2) spectra, measured by an MS instrument, are compared and
  matched against a public spectral database containing public domain
  reference MS2 spectra.

- A tutorial presenting [Use Cases and Examples for Annotation of
  Untargeted Metabolomics
  Data](https://jorainer.github.io/MetaboAnnotationTutorials/) using
  the `MetaboAnnotation` and `MetaboCoreUtils` packages
  [@Rainer:2022].

- A tutorial for [LC-MS pre-processing with
  `xcms`](https://jorainer.github.io/metabolomics2018/xcms-preprocessing.html)
  with examples on how to perform centroiding of profile-mode MS data
  in `R` and how to define data set-specific parameters for
  chromatographic peak detection, alignment and correspondence of
  untargeted (LC-MS-based) metabolomics data.


## Session information

The following packages have been used to generate this document.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown', 'msmbstyle'
), 'packages.bib')
```

```{r si}
sessionInfo()
```

## References {-}
