# Additional materials and session information {#sec-si}

## Additional materials

- The [Single-cell proteomics data analysis using `QFeatures` and
  `scp`](https://lgatto.github.io/QFeaturesScpWorkshop2021/) workshop
  is provided as two vignettes. The first one provides a general
  introduction to the `QFeatures` class in the general context of mass
  spectrometry-based proteomics data manipulation. The second vignette
  focuses on single-cell application and introduces the `scp` package
  [@Vanderaa:2021] as an extension of `QFeatures`. This second
  vignette also provides exercises that give the attendee the
  opportunity to apply the learned concepts to reproduce a published
  analysis on a subset of a real data set. A [recent
  workshop](https://github.com/lgatto/2024_scpworkshop_EUBIC), offered
  at the 2024 EuBIC winter school, provides teaching material for the
  new *scplainer* analysis workflow.

- A tutorial presenting [Use Cases and Examples for Annotation of
  Untargeted Metabolomics
  Data](https://jorainer.github.io/MetaboAnnotationTutorials/) using
  the `MetaboAnnotation` and `MetaboCoreUtils` packages
  [@Rainer:2022].

- [Exploring and analyzing LC-MS data with Spectra and
  xcms](https://jorainer.github.io/xcmsTutorials/) provides an
  overview of recent developments in Bioconductor to work with mass
  spectrometry
  ([MsExperiment](https://github.com/RforMassSpectrometry/MsExperiment),
  [Spectra](https://github.com/RforMassSpectrometry/Spectra)) and
  specifically LC-MS data ([xcms](https://github.com/sneumann/xcms))
  and walks through the preprocessing of a small data set emphasizing
  on selection of data-dependent settings for the individual
  pre-processing steps.

The [SpectraTutorials](https://jorainer.github.io/SpectraTutorials) package provides three different vignettes:

- [Seamless Integration of Mass Spectrometry Data from Different
  Sources](https://jorainer.github.io/SpectraTutorials/articles/analyzing-MS-data-from-different-sources-with-Spectra.html):
  describes import/export of MS data from/to files in different format as well
  as processing and handling of MS data with the *Spectra* package.
- [Spectra: an Expandable Infrastructure to Handle Mass Spectrometry
  Data](https://jorainer.github.io/SpectraTutorials/articles/Spectra-backends.html):
  explains the concept of backends in *Spectra*, their properties, use cases
  along with performance considerations.
- [MS/MS Spectra Matching with the MetaboAnnotation
  Package](https://jorainer.github.io/SpectraTutorials/articles/Spectra-matching-with-MetaboAnnotation.html):
  explains how the *Spectra* package can be used together with the
  `r BiocStyle::Biocpkg("MetaboAnnotation")` package in LC-MS/MS annotation
  workflows for untargeted metabolomics data.

## Compiling the book locally

To compile and render the teaching material, you will also need
the `r BiocStyle::Biocpkg("BiocStyle")` package and the (slighly
modified) [Modern Statistics for Model Biology (msmb) HTML Book
Style](https://www-huber.embl.de/users/msmith/msmbstyle/) by Mike
Smith:

```{r setup2, eval = FALSE}
BiocManager::install(c("bookdown", "BiocStyle", "lgatto/msmbstyle"))
```

Clone the [book
repository](https://github.com/Rformassspectrometry/book) and render
the book with

```{r render, eval = FALSE}
bookdown::render_book(".")
```


## Session information

The following packages have been used to generate this document.

```{r include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown', 'msmbstyle'
), 'packages.bib')
```

```{r si}
sessionInfo()
```
